---
title: "R Notebook"
output: html_notebook
---

Martina does some modelling with the lib `leaps` for linear models. I try to do something similar here for my tumors. I take one just the one example of PAAD. First of all, there are hypermutators. As I'm still not 100% sure how to handle them, I crop the distribution at a passenger number equal to the 90th percentile.

```{r}
modelling.data <- projects

for (name in projects.names){
  modelling.data[[name]]$clinical <- subset(modelling.data[[name]]$clinical, modelling.data[[name]]$clinical$passenger_freq < quantile(modelling.data[[name]]$clinical$passenger_freq, 0.9))
}
```

I wonder at other % of availability how the patient numbers decrease after removing NAs:

```{r, fig.width=9, fig.height=6}
by <- 2
steps <- seq(0, 100, by=by)
percentage.patients <- matrix(nrow = length(steps), ncol=length(projects.names))
colnames(percentage.patients) <- projects.names
rownames(percentage.patients) <- steps

get_patient_freq <- function(clinical, x){
  initial_len = length(clinical$ID)
  avail <- get_availability(clinical, x)
  clinical <- clinical %>% select(rownames(avail))
  clinical <- na.omit(clinical)
  return((length(clinical$ID)/initial_len)*100)
}

for (name in projects.names){
  for (i in steps){
    percentage.patients[i/by, name] <- get_patient_freq(projects[[name]]$clinical, i)
  }
}

getlabels <- function(x){
  if (as.numeric(x) %% 10 == 0){
    return(x)
  }else{
    return("")
  }
}

percentage.patients <- as.data.frame(percentage.patients)
percentage.patients <- rownames_to_column(percentage.patients, var="steps")
melted.percent <- melt(percentage.patients, id=c("steps"))
#melted.percent[is.na(melted.percent)] <- 100
gp <- ggplot(melted.percent, aes(x=steps, y=value, col=variable)) +
  geom_path(aes(group=1)) +
  scale_x_discrete(limits=factor(steps), labels=sapply(steps, getlabels)) +
  theme_minimal() +
  xlab("Filtering threshold for availability") +
  ylab("Percentage retained patients (after NA omission)")

gp
```
The above plot goes up since, as we remove more and more variables, the dataset is cropped less and less. It should be a 3D plot also correlating the number of variables, but I cannot really make it.
The missingness of the variables doesn't have to "match". This is why we get values lower than the filtering threshold.

I just pick an arbitrary 80% availability that seems decent for most tumours (except for KIRK? Or maybe KICH? The colours are pretty bad). I disregard all clinical variables that have less than 80% availability from the dataframe (as I crop the data at the start this means that I need to recalculate the availabilities, but I expect them to be marginally different)

```{r}
for (name in projects.names) {
  modelling.data[[name]]$availability <- get_availability(modelling.data[[name]]$clinical, cutoff = 80)
}
```

As most models don't work with missing values, i remove all NAs from the clinical frame after selecting only the columns with 80% availability that we detected before.

```{r}
for (name in projects.names) {
  modelling.data[[name]]$clinical <- modelling.data[[name]]$clinical %>% select(rownames(modelling.data[[name]]$availability))
  modelling.data[[name]]$clinical <- na.omit(modelling.data[[name]]$clinical)
}
```

Let's see some distributions now that the data was cropped and cleaned of NAs:

```{r}
for (name in projects.names) {
  plotjitter(modelling.data[[name]]$clinical, "passenger_freq", "Number of Passenger mutations", title = paste("Driver vs Passenger - filtered", name))
}
```

I'm hopeful. Now I want to see the most frequent types of clinical variables:

```{r, fig.width=12, fig.height=9}
unique.names <- c()
for (name in projects.names){
  unique.names <- unique(c(unique.names, colnames(modelling.data[[name]]$clinical)))
}
variable_presence <- matrix(nrow = length(projects.names), ncol = length(unique.names))
colnames(variable_presence) <- unique.names
rownames(variable_presence) <- projects.names
for (name in projects.names){
  variable_presence[name,] <- (unique.names %in% colnames(modelling.data[[name]]$clinical))
}
variable_presence <- 1*variable_presence
# Remove variables that are not clinical variables
variable_presence <- variable_presence[, !(colnames(variable_presence) %in% c("frequency", "passenger_freq", "ID"))]
Heatmap(
  t(variable_presence), col = c("red", "green"), name = "Present?",
  cluster_rows = FALSE, cluster_columns = FALSE,
  show_column_dend = FALSE,
  show_row_dend = FALSE,
  column_names_rot = -45,
  column_title = "Presence of clinical variables (filtered dataset, >80% available)",
  show_heatmap_legend = FALSE,
  border = TRUE
)
# Just for reference, I also do -1 and 40% availability

make_avail_heat <- function (modelling.data, cutoff, unique.names = NULL){
    for (name in projects.names) {
    modelling.data[[name]]$availability <- get_availability(modelling.data[[name]]$clinical, cutoff = cutoff)
    }
    for (name in projects.names) {
    modelling.data[[name]]$clinical <- modelling.data[[name]]$clinical %>% select(rownames(modelling.data[[name]]$availability))
    modelling.data[[name]]$clinical <- na.omit(modelling.data[[name]]$clinical)
    }
  if (is.null(unique.names)){
    unique.names <- c()
  for (name in projects.names){
    unique.names <- unique(c(unique.names, colnames(modelling.data[[name]]$clinical)))
  }
  }
  variable_presence <- matrix(nrow = length(projects.names), ncol = length(unique.names))
  colnames(variable_presence) <- unique.names
  rownames(variable_presence) <- projects.names
  for (name in projects.names){
    variable_presence[name,] <- (unique.names %in% colnames(modelling.data[[name]]$clinical))
  }
  variable_presence <- 1*variable_presence
  # Remove variables that are not clinical variables
  variable_presence <- variable_presence[, !(colnames(variable_presence) %in% c("frequency", "passenger_freq", "ID"))]
  print(Heatmap(
    t(variable_presence), col = c("red", "green"), name = "Present?",
    cluster_rows = FALSE, cluster_columns = FALSE,
    show_column_dend = FALSE,
    show_row_dend = FALSE,
    column_names_rot = -45,
    column_title = paste("Presence of clinical variables (filtered dataset, >", cutoff, "% available)", sep = ""),
    show_heatmap_legend = FALSE,
    border = TRUE
  ))
}

all_names <- c("frequency", "passenger_freq", "ID", "age_at_diagnosis", "age_at_index",
"ajcc_clinical_m", "ajcc_pathologic_n", "ajcc_pathologic_stage",
"ajcc_pathologic_t", "alcohol_history", "alcohol_intensity",
"bmi", "cigarettes_per_day", "classification_of_tumor", "days_to_birth",
"days_to_death", "days_to_diagnosis", "days_to_last_follow_up",
"days_to_last_known_disease_status", "days_to_recurrence", "ethnicity",
"gender", "height", "icd_10_code", "last_known_disease_status",
"morphology", "primary_diagnosis", "prior_malignancy", "prior_treatment",
"progression_or_recurrence", "race", "site_of_resection_or_biopsy",
"synchronous_malignancy", "tissue_or_organ_of_origin", "tumor_grade",
"tumor_stage", "vital_status", "weight", "year_of_birth", "year_of_death",
"year_of_diagnosis", "years_smoked", "ajcc_clinical_t", "ajcc_pathologic_m",
"ajcc_staging_system_edition", "pack_years_smoked", "figo_stage",
"ajcc_clinical_n", "ajcc_clinical_stage", "alcohol_days_per_week",
"alcohol_drinks_per_day", "anaplasia_present", "anaplasia_present_type",
"ann_arbor_b_symptoms", "ann_arbor_clinical_stage", "ann_arbor_extranodal_involvement",
"ann_arbor_pathologic_stage", "asbestos_exposure", "best_overall_response",
"burkitt_lymphoma_clinical_variant", "cause_of_death", "child_pugh_classification",
"circumferential_resection_margin", "coal_dust_exposure", "cog_liver_stage",
"cog_neuroblastoma_risk_group", "cog_renal_stage", "cog_rhabdomyosarcoma_risk_group",
"days_to_best_overall_response", "enneking_msts_grade", "enneking_msts_metastasis",
"enneking_msts_stage", "enneking_msts_tumor_site", "environmental_tobacco_smoke_exposure",
"esophageal_columnar_dysplasia_degree", "esophageal_columnar_metaplasia_present",
"first_symptom_prior_to_diagnosis", "gastric_esophageal_junction_involvement",
"gleason_grade_group", "goblet_cells_columnar_mucosa_present",
"gross_tumor_weight", "igcccg_stage", "inpc_grade", "inpc_histologic_group",
"inrg_stage", "inss_stage", "irs_group", "irs_stage", "ishak_fibrosis_score",
"iss_stage", "laterality", "lymph_nodes_positive", "lymph_nodes_tested",
"lymphatic_invasion_present", "masaoka_stage", "medulloblastoma_molecular_classification",
"metastasis_at_diagnosis", "metastasis_at_diagnosis_site", "method_of_diagnosis",
"micropapillary_features", "mitosis_karyorrhexis_index", "perineural_invasion_present",
"peripancreatic_lymph_nodes_positive", "peripancreatic_lymph_nodes_tested",
"premature_at_birth", "primary_gleason_grade", "radon_exposure",
"residual_disease", "respirable_crystalline_silica_exposure",
"secondary_gleason_grade", "smoking_frequency", "supratentorial_localization",
"time_between_waking_and_first_smoke", "tobacco_smoking_onset_year",
"tobacco_smoking_quit_year", "tobacco_smoking_status", "tumor_confined_to_organ_of_origin",
"tumor_focality", "tumor_largest_dimension_diameter", "tumor_regression_grade",
"type_of_smoke_exposure", "type_of_tobacco_used", "vascular_invasion_present",
"vascular_invasion_type", "weeks_gestation_at_birth", "weiss_assessment_score",
"wilms_tumor_histologic_subtype")

make_avail_heat(projects, -1, unique.names=all_names)
make_avail_heat(projects, 0, unique.names=all_names)
make_avail_heat(projects, 10, unique.names=all_names)
make_avail_heat(projects, 50, unique.names=all_names)
make_avail_heat(projects, 60, unique.names=all_names)
make_avail_heat(projects, 100, unique.names=all_names)
```

The last thing we need to do is remove all the IDs that have been shed from the clinical data in the binary presence/absence matrix, as we will need them for modelling. I also fuse them in just a single frame I name "modelling".

```{r}
for (name in projects.names){
  modelling.data[[name]]$gene_labels
  # There are some columns that are repeated in the gene labels, so I first remove them
  modelling.data[[name]]$gene_labels$frequency <- NULL
  modelling.data[[name]]$gene_labels$passenger_freq <- NULL
  modelling.data[[name]]$modelling <- merge(modelling.data[[name]]$clinical, modelling.data[[name]]$gene_labels, by="ID")
  # We could get rid of the two extra frames here, but the data is pretty small so I don't bother.
}
```

There are various factors as clinical variables. I'm curious about the meaning of the ajcc pathologic stage. It's also known as the TNM staging system.

The ajcc pathologic states describe the invasiveness of the tumor:
* T - referring to the primary tumor mass, measures its extent:
  + TX: Cannot be measured
  + T0: Cannot be found
  + T1-4: The larger the cell mass, the larger the number
* N - referring to regional lymph nodes:
  + NX: Cannot be measured
  + N0: There is no cancer in nearby lymph nodes
  + N1-3: Number and location of lymph nodes that show the presence of cancer cells.
* M - Referring to the presence of metastases:
  + MX: Cannot be measured
  + M0: Cancer has not spread to other parts of the body
  + M1: Cancer has spread to other parts of the body

(Source for the above: NIH)

Martina evaluated the presence or absence of a mutation in the model as variables. However, it doesn't sound very good to me. We are including as variables in the model the very thing that we are modelling...!

I'll just do one model for PAAD to see how the models could look like. I'll use the obs/10 rule to select the number of observations. I'll also just keep the top 20 most-frequent mutations (at the gene level) as the only descriptors, although they will be fewer than that, seeing the number of drivers.

```{r}
PAAD <- modelling.data$`TCGA-PAAD`
nr.obs <- round(length(PAAD$clinical$ID)/10)
# We get 14 possible variables

# Get the 20 most-frequently mutated genes
genefreqs <- lapply(PAAD$gene_labels[!(names(PAAD$gene_labels) %in% c("frequency", "passenger_freq", "ID"))], sum)
wrap <- function(x){grepl(pattern="_", x=x, fixed=TRUE)}
genefreqs <- genefreqs[!(unlist(lapply(names(genefreqs), wrap)))]
top20genes <- names(sort(unlist(genefreqs), decreasing = TRUE)[1:min(length(genefreqs),20)])
# There are only 8 mutated genes in PAAD, so there's that.

# extract the model data
model_data <- PAAD$modelling
model_data <- select(model_data, c(names(PAAD$clinical), all_of(top20genes)))

# We need to make the ajcc levels as binary variables.
####################################### HUGE NOTE ######################################################
########### I make the choice here to say that "cannot determine stage" is equal to 0 ##################
########### This is 100% not true!!! It should become an NA, but this way like, almost #################
########### all patients are removed due to the presence of NAs. What do we do?? #######################
########################################################################################################
model_data$ajcc_pathologic_m <- revalue(model_data$ajcc_pathologic_m, c("M0" = 0, "MX" = 0, "M1" = 1))
model_data$ajcc_pathologic_n <- revalue(model_data$ajcc_pathologic_n, c("N0" = 0, "NX" = 0, "N1" = 1, "N2" = 1, "N3" = 1, "N1b" = 1))

# We can get rid of the ID
model_data$ID <- NULL

# The model fails if there are factors with just one level. So I remove factors with just the one level.
get_nlevels <- function(x){
  if (is.factor(x)){
    return(nlevels(x))
  }else{
    return(Inf)
  }
}

model_data <- model_data[, sapply(model_data, get_nlevels) > 1]
```
We can now use `regsubsets` in order to find the best variables to include in the model. First, we do the full models and take a look at them:

```{r}
full.model.drivers <- lm(frequency ~ ., model_data)
full.model.passeng <- lm(passenger_freq ~ ., model_data)
full.model.both <- lm(cbind(frequency, passenger_freq) ~ ., model_data)
summary(full.model.drivers)
summary(full.model.passeng)
summary(full.model.both)
```


Since the things below failed, I'll just make the reduced models here:
```{r}
reduced.model.drivers <- lm(frequency ~ ajcc_pathologic_n+KRAS+TP53+SMAD4+GNAS+CDKN2A+KDM6A+RNF43+U2AF1, model_data)
reduced.model.passeng <- lm(passenger_freq ~ ajcc_pathologic_n+KRAS+TP53+SMAD4+GNAS+CDKN2A+KDM6A+RNF43+U2AF1, model_data)
reduced.model.both <- lm(cbind(frequency, passenger_freq) ~ ajcc_pathologic_n+KRAS+TP53+SMAD4+GNAS+CDKN2A+KDM6A+RNF43+U2AF1, model_data)

summary(reduced.model.drivers)
summary(reduced.model.passeng)
summary(reduced.model.both)
```

## Not complete yet

Now we can do the regression, and keep only the most important variables to keep:

```{r}
# Exhaustive search cannot handle this many variables. so I'm limiting them to just the nodal status and the
# gene labels
best.models.drivers <- regsubsets(frequency ~ ajcc_pathologic_n+KRAS+TP53+SMAD4+GNAS+CDKN2A+KDM6A+RNF43+U2AF1, model_data)
best.models.passeng <- regsubsets(passenger_freq ~ ajcc_pathologic_n+KRAS+TP53+SMAD4+GNAS+CDKN2A+KDM6A+RNF43+U2AF1, model_data)

summary(best.models.drivers)
summary(best.models.passeng)
```
We now make the model, and run some model quality statistics.


```{r}
make_binary_matrix <- function(project){

}
genefreqs <- lapply(PAAD$gene_labels[!(names(PAAD$gene_labels) %in% c("frequency", "passenger_freq", "ID"))], sum)
wrap <- function(x){grepl(pattern="_", x=x, fixed=TRUE)}
genefreqs <- genefreqs[!(unlist(lapply(names(genefreqs), wrap)))]
top20genes <- names(sort(unlist(genefreqs), decreasing = TRUE)[1:min(length(genefreqs),20)])

model_data <- PAAD$clinical
model_data <- merge(model_data, PAAD$gene_labels[c(top20genes, "ID")], by="ID", all.x = TRUE)

model_data$ajcc_pathologic_m <- revalue(model_data$ajcc_pathologic_m, c("M0" = 0, "MX" = NA, "M1" = 1))
model_data$ajcc_pathologic_n <- revalue(model_data$ajcc_pathologic_n, c("N0" = 0, "NX" = NA, "N1" = 1, "N2" = 1, "N3" = 1, "N1b" = 1))

# We can get rid of the ID
model_data$ID <- NULL

# Remove the variables with less than 80% availability
model_data <- model_data[rownames(get_availability(model_data, 80))]
model_data <- na.omit(model_data)

# The model fails if there are factors with just one level. So I remove factors with just the one level.
get_nlevels <- function(x){
  if (is.factor(x)){
    return(nlevels(x))
  }else{
    return(Inf)
  }
}

model_data <- model_data[, sapply(model_data, get_nlevels) > 1]

```

Let's say I keep the top 20 most mutated *genes* and consider their binary presence (0 for "not mutated" and 1 for "mutated") in the model. For most tumours that have very few numbers of drivers (like PAAD), this will be all genes. I also transpose the N and M metrics as MX -> NA, M0 -> 0 and M1 -> 1, and NX -> NA, N0 -> 0 and N1-3 -> 1.
