---
title: "Misc Analysis"
output: html_notebook
---

This file will be a acatch-all for all the small things that are not worth a separate analysis for.

# Preamble

As per usual, we begin by loading some libraries.

```{r global options, include = FALSE}
# Do not include warnings in the output PDF
knitr::opts_chunk$set(warning=FALSE)
```

```{r, echo=TRUE, results='hide'}
library(plyr) # Data shape casting (needs to be loaded before dplyr)
library(tidyverse)
# Tidyverse loads ggplot2, purrr, tibble, dplyr, tidyr, stringr, readr and forcats
# This suppresses the warning that summarise makes about grouping.
options(dplyr.summarise.inform = FALSE)
library(reshape2) # Data melt and cast
library(cowplot) # Margin plots made (a bit) easier
library(gridExtra) # Stuff like grid.arrange

# The docstring package is entirely optional but it allows me to print out help messages
# as I write more code, so it's nice as I often forget what my own functions do...
library(docstring)
```

Other packages that need to be installed (I call them sporadically without bringing them into scope):
```{r}
test_installed <- function (package) {
  if(!package %in% installed.packages()) {
    stop("Package '", package, "' is not installed.")
  }
}
test_installed("ggrepel") # Add labels over scatterplots
test_installed("jtools") # Provides easy effect plots
```


# Loading data

Data preprocessing has already occurred. Please take a look at the `chasm_data_preprocessing.R` script in the GitHub. I only load the smaller dataset, with just the five tumor types of interest.

```{r}
load("F:/Data/University/Thesis/Data/CHASMplus/preprocessed_CHASM_data_2.Rdata")
p.names <- names(projects)

# I save the tumor type in the clinical data as an attribute, as basically every function
# that takes the clinical data also needs it (like, for plotting)
# This way, I don't need to pass it manually every time, but just extract it from the data
# itself.
for (name in p.names) {
  attr(projects[[name]]$clinical, "project.id") <- name
}
```


# ParsSNP missing annotations

```{r}
# I'm loading the test file from ParsSNP

pars.annotated <- read.table(
  "F:/Data/University/Thesis/Tools/ParsSNP/tcga_annotated.hg19_multianno.txt",
  header = TRUE,
  sep = "\t"
  )

# Removing the columns that are not descriptors

pars.annotated.pred <- pars.annotated[, -which(names(pars.annotated) %in% c("Chr", "Start", "End", "Ref", "Alt"))]

tot.nas <- sum(is.na(pars.annotated.pred))
tot.cells <- length(pars.annotated.pred) * length(pars.annotated.pred[, 1])

tot.nas/tot.cells*100

```


# CHASMplus enrichment gene lists

```{r}
unique(projects$`TCGA-BRCA`$crystal$Hugo[projects$`TCGA-BRCA`$crystal$is.driver]) -> BRCAdrivers

unique(projects$`TCGA-COAD`$crystal$Hugo[projects$`TCGA-COAD`$crystal$is.driver]) -> COADdrivers

unique(projects$`TCGA-LUAD`$crystal$Hugo[projects$`TCGA-LUAD`$crystal$is.driver]) -> LUADdrivers

unique(projects$`TCGA-SKCM`$crystal$Hugo[projects$`TCGA-SKCM`$crystal$is.driver]) -> SKCMdrivers

unique(projects$`TCGA-GBM`$crystal$Hugo[projects$`TCGA-GBM`$crystal$is.driver]) -> GBMdrivers

# Just do ...drivers %>% paste(collapse = "\n) %>% cat to get the gene lists that can be pasted to EnricR

```


# Count of all passenger genes across all projects

```{r}
map(projects, function(x) {
  return(x$crystal$Hugo[!x$crystal$is.driver])
}) %>%
  reduce(c) %>%
  unique() %>%
  length() -> pass.count

# This is taken from the upset plot
11271 / pass.count # Five tumor types
sum(11271, 31, 215, 375, 1289, 2076) / pass.count # Four or more
```


Novel driver genes per tumor type

```{r}
cgc.genes <- readLines("F:/Data/University/Thesis/Data/UniqueCosmicMutations.txt")

novel <- list()
novel$BRCA <- BRCAdrivers[!BRCAdrivers %in% cgc.genes]
novel$LUAD <- LUADdrivers[!LUADdrivers %in% cgc.genes]
novel$COAD <- COADdrivers[!COADdrivers %in% cgc.genes]
novel$GBM <- GBMdrivers[!GBMdrivers %in% cgc.genes]
novel$SKCM <- SKCMdrivers[!SKCMdrivers %in% cgc.genes]

print(novel)
```
